Querying Integrated Scientific Observation and Measurement Data

### 2010-06-07 docs/algdraft/DataIntegration/method.tex
update the algorithm with a framework (to finish)

As we analyzed in Section \ref{sec:relatedwork}, there are two extrems
in querying integrated data. 
One is to rewrite the original query to a series of new queries over
the data; the other is to materialize the data to a consistent data
model and then answer the query over the materialized database. 
In what follows we work from these two directions to leverage the
semantic information in answering such queries.

@@

%Framework of query rewriting
Roughly speaking, the query rewriting consists of two steps:
\begin{itemize}
\item From the given query, search the database metadata and
  annotation information to find the relevant data tables and
  attributes that need to be used to answer the given query. 
\item Translate the given OM query to queries over the relevant data
  tables. 
\end{itemize}

@@

%Elaborate the first step
The first step is to map the OM concepts in the query to the real data
structure. 
With the annotation structures $OT$, $CT$, and $MT$, any given OM
concepts can be linked to a measurement type $mt_{id}$. 
Then, the annotation structure $Map$ can be used to bridge OM concepts and data sets. 
Utilizing $Map$, we can find the attributes $attr_i$ and the
annotation $A_{id}$).
E.g., when the given concept is {\em measurement}, we can directly get
its related attribute names and annotation id $A_{id}$. 
In case that the given OM concept is a non-measurement concept, we can use
other annotation structures to figure out the measurementt types. 
E.g., if the OM concept is {\em characteristic} ``Species'', we can use
$MT$ to get the measurement type id $mt_{id}$. 
After we get $A_{id}$, we can quickly find the dataset id $d_{id}$
using $AD$ metadata table (Section \ref{sec:annotation}). 
Then, we get the relevant data tables and the needed attributes.

%Elaborate the second step
Let $D_{cand}$ be the set of relevant (or candidate) data tables and $d_{a_1} \cdots d_{a_m}$
are the related attributes for table $d\in D_{cand}$. 
The second step translates the original OM query to SQL queries over the data tables.

@@

>>>>

As we analyzed in Section \ref{sec:relatedwork}, there are two extremes
in querying integrated data.
One is to rewrite the original query to a series of new queries over
the data; the other is to materialize the data to a consistent data
model and then answer the query over the materialized database.
In what follows we work from these two directions to leverage the
semantic information in answering such queries.

@@

%Framework of query rewriting
Roughly speaking, the query rewriting consists of two steps:
\begin{itemize}
\item From the given query with entity, characteristic information, search the database metadata and
  annotation information to find the relevant data tables and
  attributes that need to be used to answer the given query.
\item Translate the given OM query to queries over the relevant data
  tables. To do this translation, we decompose an OM query to have several basic component operators. 
  \begin{itemize}
  \item Basic OM query with one entity type. While the characteristic condition may be conjunctive or disjunctive, it may also has aggregation. E.g., $A(Cha1>25.0 && Count(Cha)>3, Height>2.0)$
  \item Context OM chain refer to a set of queries having ``HasContext'' relationship. Obviously, these queries have different entity condition. E.g., $Ent1(...) --> Ent2(), Ent1()-->Ent3()$
  \item Disjunctive OM query with multiple entity types. $Ent1(...), Ent2(), Ent3(...)$.
  \end{itemize}

\end{itemize}

@@

%Elaborate the first step
The first step is to map the OM concepts in the query to the real data
structure.
With the annotation structures $OT$, $CT$, and $MT$, any given OM
concepts can be linked to a measurement type $mt_{id}$.
Then, the annotation structure $Map$ can be used to bridge OM concepts and data sets.
Utilizing $Map$, we can find the attributes $attr_i$ and the
annotation $A_{id}$).
E.g., when the given concept is {\em measurement}, we can directly get
its related attribute names and annotation id $A_{id}$.
In case that the given OM concept is a non-measurement concept, we can use
other annotation structures to figure out the measurement types.
E.g., if the OM concept is {\em characteristic} ``Species'', we can use
$MT$ to get the measurement type id $mt_{id}$.
After we get $A_{id}$, we can quickly find the dataset id $d_{id}$
using $AD$ metadata table (Section \ref{sec:annotation}).
Then, we get the relevant data tables and the needed attributes.

%Elaborate the second step
Let $D_{cand}$ be the set of relevant (or candidate) data tables and $d_{a_1} \cdots d_{a_m}$
are the related attributes for table $d\in D_{cand}$.

The second step translates the original OM query to SQL queries over the data tables.
From now on, we assume that we know the attribute already for all the needed characteristics.

To translate the given query:
 Group the basic queries (with only one entity type) to list of context chains.

\begin{algorithm}[htb]
\begin{algorithmic}[1]
%\begin{footnotesize}
\STATE $Result$ := $\emptyset$; 
\COMMENT{//From the given query, get the DNF of context queries.}
\STATE $contextChainDNF$ := ${\bf GroupBasicQuery}(OMQ)$;
\STATE
\COMMENT{//Execute every DNF int the context chain set and union their results}
\FOR{each context chain $oneContextQuery$ in $contextChainDNF$}
	\STATE $result_{CNF}$ := $\emptyset$;
    \COMMENT{//Execute every basic query in this CNF query and intersect their results}
    \FOR{each basic query $OMBasic$ in $ContextQuery$}
    \STATE $result_{basic}$ := $OMBasic.execute()$;
    \STATE $result_{CNF}$ := $result_{CNF} \cap result_{basic}$;
    \ENDFOR
	\STATE $Result$ := $Result\cup result_{CNF}\$
\ENDFOR
\RETURN $RESULT$;
%\end{footnotesize}
\end{algorithmic}
\caption{Execute(OMQ)}
\label{alg:omq}
\end{algorithm}

\begin{figure}[htb]
%\normalsize
%\begin{minipage}{5in}
%\sf
\begin{scriptsize}
\begin{tabbing}
1166\= 16 \= 16 \= 16 \= 16 \= 16 \= 16 \= 16 \= 16 \= \kill
{\bf Algorithm GroupBasicQuery}($OMQ$);\\
1).\>$contextChainSet$ := $\emptyset$;\\
\\
\>//Put the basic queries in context into context chains
2).\>{\bf for} (each basic context $OMBasic-->ContextOMBasic$ in $OMQ$)\\
3).\>\>$added$ := false;
4).\>\>{\bf for} (each set $oneContextChain \in contextChainDNF)\\
5).\>\>\>{\bf if} (($OMBasic \in oneContextChain$) or $ContextOMBasic \in oneContextChain$))\\
6).\>\>\>\> $oneContextChain$ = $oneContextChain \cup OMBasic \cup ContextOMBasic$\\
7).\>\>\>\> $added$ :=true;\\
8).\>\>\>\> {\bf break;}\\
9).\>\>{\bf if}($added$==false)\\
10).\>\>\>Create a set $newContextChain$ with two elements $OMBasic$ and $ContextOMBasic$;\\
11).\>\>\>Add $newContextChain$ to $contextChainSet$;\\
\\
\>//Put the basic queries in context into context chains
12).\>{\bf for} (each $OMBasic$ not in context of $OMQ$)\\
13).\>\>Create a set $newContextChain$ with one element $OMBasic$\\
14).\>\>Add $newContextChain$ to $contextChainSet$;\\
15).\>{\bf return} $contextChainSet$;\\
\end{tabbing}
\end{scriptsize}
%\end{minipage}
\caption{GroupBasicQuery(OMQ)}
\label{alg:omq}
\end{figure}

@@

First, for the basic OM query with one entity type.

Second, for the context chain query.
